---
title: "Geography 176A"
author: '[Kyler Plouffe](https://kplouffe.github.io/kplouffe/)'
subtitle: 'Lab 05: Rasters and Remote Sensing'
output:
  html_document:
    theme: journal
  pdf_document: default
---


```{r setup, include=FALSE}
# SPDS
library(tidyverse)
library(sf)
library(units)

# Data
library(USAboundaries)
library(rnaturalearth)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)
library(rmapshaper)
library(mapview)
library(leaflet)
library(elevatr)
library(raster)
library(getlandsat)

knitr::opts_chunk$set(cache = F)
```

### Question 1
```{r}
setwd("C:/Users/slaye/Documents/github/geog-176A-labs/")
bb=read_csv("data/uscities.csv") %>% 
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>% 
  filter(city == "Palo", state_name == "Iowa") %>% 
  st_transform(5070) %>% 
  st_buffer(5000) %>% 
  st_bbox() %>% 
  st_as_sfc() %>% 
  st_as_sf()
```

### Question 2
```{r}
#2.2
setwd("C:/Users/slaye/Documents/github/geog-176A-labs/")
meta = read_csv("data/palo-flood-scene.csv")

files = lsat_scene_files(meta$download_url) %>% 
  filter(grepl(paste0("B", 1:6, ".TIF$", collapse = "|"), file)) %>% 
  arrange(file) %>% 
  pull(file)

#2.3
st = sapply(files, lsat_image)

s = stack(st) %>% setNames(c(paste0("band", 1:6)))

#2.4
cropper = bb %>% st_transform(crs(s))

r = crop(s, cropper)

r
```
dimensions : 340, 346, 117640, 6  (nrow, ncol, ncell, nlayers)
crs        : +proj=utm +zone=15 +datum=WGS84 +units=m +no_defs 
resolution : 30, 30  (x, y)

### Question 3
```{r}
#1.1
par(mfrow = c(2,2))
RGB = plotRGB(r, r=4, g=3, b=2)
NIRRG = plotRGB(r, r=5, g=4, b=3)
NIRSWIR1G = plotRGB(r, r=5, g=6, b=4)
TIRSCoastalCirrus = plotRGB(r, r=10, g=1, b=9)

#3.2
par(mfrow = c(2,2))
RGB2 = plotRGB(r, r=4, g=3, b=2, stretch = "lin")
NIRRG2 = plotRGB(r, r=5, g=4, b=3, stretch = "lin")
NIRSWIR1G2 = plotRGB(r, r=5, g=6, b=4, stretch = "hist")
TIRSCoastalCirrus2 = plotRGB(r, r=10, g=1, b=9, stretch = "hist")

```
The stretch improves the contrast in th emiage so you can see the color differences better.

### Question 4
```{r}
#4.1
ndvi = (r$band5 - r$band4) / (r$band5 + r$band4)
ndwi = (r$band3 - r$band5) / (r$band3 + r$band5)
mndwi = (r$band3 - r$band6) / (r$band3 + r$band6)
wri = (r$band3 + r$band4) / (r$band5 + r$band6)
swi = 1 / sqrt(r$band2 - r$band6)

palette = colorRampPalette(c("blue", "white", "red"))

(water = stack(ndvi, ndwi, mndwi, wri, swi) %>% 
  setNames(c("vegetation", "water", "modwater", "waterratio", "simplewater")))
plot(water, col = palette(256))

#4.2
thresholding1 = function(x){ifelse(x < 0, 1, NA)}
thresholding2 = function(x){ifelse(x > 0, 1, NA)}
thresholding3 = function(x){ifelse(x > 0, 1, NA)}
thresholding4 = function(x){ifelse(x > 1, 1, NA)}
thresholding5 = function(x){ifelse(x < 5, 1, NA)}

f1 = calc(ndvi, thresholding1)
f2 = calc(ndwi, thresholding2)
f3 = calc(mndwi, thresholding3)
f4 = calc(wri, thresholding4)
f5 = calc(swi, thresholding5)

(flood = stack(f1, f2, f3, f4, f5) %>% 
  setNames(c("vegetation", "water", "modwater", "waterratio", "simplewater")))
plot(flood, col = "blue")



```

### Question 5
```{r}
#5.1
set.seed

#5.2
rvalues = getValues(r) %>% 
  na.omit
dim(rvalues)
#these values indicate that the values were extracted as points from 2 separate groups
par(mfrow = c(2,2))
kave = stats::kmeans(rvalues, 4)
kmeans_raster = flood$water
values(kmeans_raster) = kave$cluster
plot(kmeans_raster)

kave = stats::kmeans(rvalues, 8)
kmeans_raster = flood$water
values(kmeans_raster) = kave$cluster
plot(kmeans_raster)

kave = stats::kmeans(rvalues, 12)
kmeans_raster = flood$water
values(kmeans_raster) = kave$cluster
plot(kmeans_raster)

kave = stats::kmeans(rvalues, 16)
kmeans_raster = flood$water
values(kmeans_raster) = kave$cluster
plot(kmeans_raster)

```
```{r}
#5.3
kave = stats::kmeans(rvalues, 14)
kmeans_raster = flood$water
values(kmeans_raster) = kave$cluster
plot(kmeans_raster)

f1[is.na(f2[])] <- 0 
f2[is.na(f2[])] <- 0 
f3[is.na(f2[])] <- 0 
f4[is.na(f2[])] <- 0 
f5[is.na(f2[])] <- 0 

floodtable = table(values(f2), values(kmeans_raster))
floodtable
which.max(floodtable)

thresholding6 = function(x){ifelse(x > 5, 1, NA)}
f6 = calc(kmeans_raster, thresholding6)

flood2 = stack(f1, f2, f3, f4, f5, f6) %>% 
  setNames(c("vegetation", "water", "modwater", "waterratio", "simplewater", "kmeans"))
plot(flood2, col = "blue")

```

### Question 6
```{r}
stats = cellStats(flood2, sum)
stats = (stats*900)/1000
knitr::kable(stats, 
             caption = "Area in km^2 Affected by Flooding per Raster",
            
             format.args = list(big.mark = ","))

uncertain = calc(flood2, fun = sum)

plot(uncertain)
mapview(uncertain)

```
Some of the cell values are not an even number because the data is averaged over the raster box, meaning the boxes were partially overlapped somehow which leads to a number that is not an integer.

### Extra Credit
```{r error = TRUE}
bb2 = st_transform(bb, 4326)

osm = osmdata::opq(bb2) %>% 
  osmdata::add_osm_feature("building") %>% 
  osmdata::osmdata_sf()
leaflet() %>% 
  setView(lng=-91.78943, lat=42.06307, zoom = 13) %>% 
  addProviderTiles(providers$CartoDB) 

local <- st_point(c(-91.78943, 42.06307))

sf_local = st_sfc(local, crs = st_crs(flood2)) %>% 
  st_cast("POINT")

sf_local

floodvalue = raster::extract(flood2, sf_local, na.rm = TRUE)

flood2
```







