---
title: "Geography 176A"
author: "[Kyler Plouffe](https://kplouffe.github.io/kplouffe/)"
subtitle: 'Lab 03: COVID-19 Pandemic'
output:
  html_document:
    theme: journal
---
## Question 1
```{r}
#attach packages
library(tidyverse)
library(knitr)
library(readxl)
library(zoo)

covid = read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')

head(covid, 5)

countydata = covid %>% 
  filter(state == "California") %>% 
  group_by(county) %>% 
  mutate(newCases = cases - lag(cases)) %>% 
  ungroup()
 
```

 
```{r}
countydata2= countydata %>% 
  filter(date == max(date))

(most_new_cases = countydata2 %>% 
  slice_max(newCases, n = 5) %>% 
  select(county, newCases))
  
(most_cumulative_cases = countydata2 %>% 
  slice_max(cases, n = 5) %>% 
  select(county, cases))

knitr::kable(most_new_cases, 
             caption = "Most New Cases California Counties",
             col.names = c("County", "New Cases"),
             format.args = list(big.mark = ","))
knitr::kable(most_cumulative_cases, 
             caption = "Most Cumulative Cases California Counties",
             col.names = c("County", "Total Cases"),
             format.args = list(big.mark = ","))
```
  

```{r}
library(readxl)
pop <- read_excel("~/github/geog-176A-labs/data/PopulationEstimates.xls", skip=2)
names(pop)

p2=pop %>% 
  select(fips=FIPStxt, state= State, county=Area_Name, pop2019=POP_ESTIMATE_2019) %>% 
  filter(state=="CA") %>% 
  group_by(county) %>% 
  slice_max(pop2019, n=1) %>% 
  select(fips, pop2019)
  
covidpop = inner_join(countydata, p2, by="fips")
count(covidpop)
head(covidpop)
```

```{r}
covidpop2=covidpop %>% 
  filter(date==max(date))

(most_new_cases_capita = covidpop2 %>% 
  mutate(CPC=cases/pop2019) %>% 
  slice_max(CPC, n = 5) %>% 
  select(county.x, CPC))
  
(most_cumulative_cases_capita = covidpop2 %>% 
  mutate(NCPC=newCases/pop2019) %>% 
  slice_max(NCPC, n = 5) %>% 
  select(county.x, NCPC))

knitr::kable(most_new_cases_capita, 
             caption = "Most New Cases per Capita in California Counties",
             col.names = c("County", "New Cases per Capita"),
             format.args = list(big.mark = ","))
knitr::kable(most_cumulative_cases_capita, 
             caption = "Most Cumulative Cases per Capita in California Counties",
             col.names = c("County", "Total Cases per Capita"),
             format.args = list(big.mark = ","))
```

```{r}
library(dplyr)
covidpop3=covidpop %>% 
  filter(date >= max(date)-14) %>% 
  group_by(county.x, fips) %>% 
  summarize(newCases=sum(newCases)) %>% 
  inner_join(p2, by="fips") %>% 
  select(county,newCases,pop2019)
print((sum(covidpop3$newCases)/sum(covidpop3$pop2019))*100000)


```
The total number of new cases for the state of California is 
```{r}
print(sum(most_cumulative_cases$cases))
```
The total number of new cases for the state of California is
```{r}
print(sum(most_new_cases$newCases))
```
The safest counties that have not yet broken the COVID-19 guidelines include:
```{r}
safe_county=covidpop3 %>%
  mutate(guideline=(newCases/pop2019)*100000) %>% 
  filter(guideline<100) %>% 
  pull(county) ->
  safe_counties
  
print(safe_counties)
```

## Question 2
```{r}
state.of.interest=c("New York", "California", "Louisiana", "Florida")
covid %>%
  filter(state==state.of.interest) %>%
  group_by(state, date) %>%
  summarise(cases = sum(cases)) %>%
  ungroup() %>%
  mutate(newCases = cases - lag(cases),
         roll7 = rollmean(newCases, 7, fill = NA, align="right")) %>%
  ggplot(aes(x = date, y = newCases, color = state)) +
  geom_line(size = 1) +
  facet_wrap(~state) +
  ggthemes::theme_solarized() +
  theme(legend.position = 'NA') +
  labs(title = "Daily New Case Counts",
       subtitle = "Data Source: NY-Times",
       x = "Date",
       y = "New Cases")

```

```{r}
state.of.interest=c("New York", "California", "Louisiana", "Florida")
p3=pop %>% 
  select(fips=FIPStxt, state=Area_Name, pop2019=POP_ESTIMATE_2019) %>% 
  group_by(state) %>% 
  slice_max(pop2019, n=1) %>% 
  select(fips, pop2019)

percapdata=covid %>%
  filter(state %in% c("New York", "California", "Louisiana", "Florida" )) %>%
  group_by(state, date) %>% 
  summarize(cases=sum(cases), .groups="drop") %>% 
  mutate(newCases = cases - lag(cases)) %>% 
  group_by(state) %>% 
  right_join(p3, by="state") %>% 
  mutate(percap= newCases/pop2019,
         roll7 = rollmean(percap, 7, fill = NA, align="right"))
#making the second per capita graph
percapdata %>% 
  ggplot(aes(x = date, y = percap, color = state)) +
  geom_line(size = 1) +
  facet_wrap(~state) +
  ggthemes::theme_solarized() +
  theme(legend.position = 'NA') +
  labs(title = "Daily New Case Counts",
       subtitle = "Data Source: NY-Times",
       x = "Date",
       y = "New Cases per Capita")
```


Scaling by population makes some states appear much better, and makes others appear worse. This is because it places the new case counts in a ratio with the total population, so we get a better idea of what percentage of the population actually has covid.










